---
title: "Data 607 Assignment 5"
author: "Dave Greer"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
always_allow_html: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("RODBC")
library(RODBC)
library(DT)
library(ggstatsplot)
library(tidyverse)
```
## Introduction
For this assignment I decided to build a normalized database to contain all the values. I performed a basic comparison between two airlines, AM West and Alaska.  The data was derived from the assignment pdf.

```{r }
db_conn <-odbcConnect("Assignment5", rows_at_time =1)
if(db_conn == -1){
  quit("no", 1)
}
print(db_conn)
```
## Here I create the connection between my database and my R markdown sheet.

```{r  echo=FALSE}
sql <-"SET NOCOUNT ON; Select * from AirlineFlights;"
df_tstsql <- sqlQuery(db_conn, sql)
odbcClose(db_conn)
```
Rather than create a query to pull the data from the individual tables I
wrote a database view and queried the view. In this case the view is named 
AirlineFlights.

```{r}
datatable(df_tstsql)
```
I selected the fields required for analysis and I built an plot show the on-time disparity 
between AM West and Alaska.
```{r}
select(df_tstsql,Status,NumberOfMinutes,AirlineName)
dfOntime<-filter(df_tstsql,Status=="ontime")
ggplot(dfOntime, aes(x=CityName, y=NumberOfMinutes/1000))+
  geom_point(aes(color=AirlineName))

```
##Summary of Alaska On-time Minutes to Selected Cities 
```{r}
dfOntimeAlaska<-filter(df_tstsql,Status=="ontime" & AirlineName== "Alaska" )

dfNumberOfOntimeAlaskaMinutes<-select(dfOntimeAlaska,NumberOfMinutes)

summary(dfNumberOfOntimeAlaskaMinutes)
```
##Summary of AM West On-time Minutes to Selected Cities 
```{r}
dfOntimeAmWest<-filter(df_tstsql,Status=="ontime" & AirlineName== "AM West" )
dfNumberOfOntimeAMWestMinutes<-select(dfOntimeAmWest,NumberOfMinutes)
summary(dfNumberOfOntimeAMWestMinutes)
```
##Below is the plot showing the delay disparity between AM West and Alaska Airlines
```{r}
select(df_tstsql,Status,NumberOfMinutes,AirlineName)
dfDelayed<-filter(df_tstsql,Status=="delayed")
ggplot(dfDelayed, aes(x=CityName, y=NumberOfMinutes/100))+geom_point(aes(color=AirlineName))

```
##Summary of Alaska Delayed Minutes to Selected Cities 
```{r}
dfDelayedAlaska<-filter(df_tstsql,Status=="delayed" & AirlineName== "Alaska" )
dfNumberOfdfDelayedAlaskaMinutes<-select(dfDelayedAlaska,NumberOfMinutes)
summary(dfNumberOfdfDelayedAlaskaMinutes)
```
##Summary of AM West Delayed Minutes to Selected Cities 
```{r}
dfDelayedAmWest<-filter(df_tstsql,Status=="delayed" & AirlineName== "AM West" )
dfNumberOfdfDelayedAmWestMinutes<-select(dfDelayedAmWest,NumberOfMinutes)
summary(dfNumberOfdfDelayedAmWestMinutes)
```
##Analysis

##Arrivals
Based on the summary and the chart, the mean and max of on time arrival
is greater with AM West. This is especially true with regards to Phoenix.
Phoenix appears to be an outlier. Otherwise, the median arrivals is slightly 
greater with Alaska.  

##Delays
Based on the summary and the chart, the mean and max of on time delays
is greater with AM West, again this is especially true with regards to Phoenix.
Phoenix appears to be an outlier. But in the case of delays Alaska's median
is lower than AM West.




