---
title: "Homework1"
author: "Dave Greer"
date: "2023-09-09"
output: html_document
---

``````{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
##install.packages("utf8")
##install.packages("ggstatsplot")

library(here)
library(readr)
library(ggstatsplot)
library(purrr)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(qqplotr)
library(jtools) 
library(DT)
options(scipen=999)

```

## Introduction

For this assignment I decided to learn something about extramarial affairs. I 
wanted to know which gender was more likely to have an extramatrial affair.
I found a data set called "Fair's Extramarital Affairs Data".  This data was 
collected in 1969 and contains 601 observations of 9 variables about how many
extramartial affairs people had over a 12 month period. 
The variables in the dataset include age, gender, education, occupation, 
religiousness, and children. Personally I approached this assignment with a 
bias towards men have more extramartial affairs than women.

## Sources
 [Click here to view all data  sources](https://vincentarelbundock.github.io/Rdatasets/articles/data.html) ||
 [Click here to view Data Source Description](https://vincentarelbundock.github.io/Rdatasets/doc/AER/Affairs.html)
 
## Data Gathering, Collation and Simple Bar Chart
In the code chunk below I am gathering the data from a remote data source (line 45)
In lines 48 through 57 I am giving the columns in dataframe meaningful titles.
I am converting the Number of Years column to Number of Months because its easier to work with integers. On line 63 I am removing the observation that did not contain at least 1 affair.
The data is displayed in tabular format on line 65 and a simple bar chart is created on lines 67 and 68.


```{r}

dat <- read_csv('https://raw.githubusercontent.com/greerda/Data607/main/Affairs.csv')

df_affairs <- dat|> select(`Number of Affairs` = affairs,
                            Age = age,
                            Gender =gender,
                           `Months Married`= yearsmarried,
                           `Has Children` = children,
                           `Religiosity Level (1-5)`= religiousness,
                            `Highest Grade Level Achieved` =education,
                           `Level of Satisfaction with Marriage (1-5)`= rating)

df_affairs$`Months Married`<- round(df_affairs$`Months Married`*12,0)

df_affairs<-df_affairs|>filter(`Number of Affairs`>0)

datatable(df_affairs)

ggplot(df_affairs, aes(x=Gender, y=`Number of Affairs`, fill=Gender) )+
          geom_bar(stat = "identity" )+theme_minimal()+ggtitle("Number of Extra Martial Affairs by Gender") 


                      
```

## Conclusion and Recommendations 
After performing this basic analysis I was surprised that my personal biases
were overturned. I initially thought that men would have a exponentially greater
number of extramartial affairs.  Based on this information I recommend that
men should buy their wives flowers periodically when they go to Costco.

