---
title: "Sentiment Analysis"
author: "Dave Greer and Kristin Lussi"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(tidyr)
library(tidytext)
library(textdata)
library(janeaustenr)
library(dplyr)
library(ggplot2)
library(stringr)
library(wordcloud)
library(reshape2)
library(tidyverse)
library(httr)
library(jsonlite)
library(SentimentAnalysis)
```

### Get Sentiments for AFINN
```{r}
get_sentiments("afinn")
```
### Get Sentiments for Bing
```{r}
get_sentiments("bing")
```

### Get Sentiments for NRC
```{r}
get_sentiments("nrc")
```
### Creating A tibble of books
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

```
### Creating a tibble of positive sentiments using NRC and the book Emma
```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```
### Creating a tibble of sentiments using Bing   
```{r}

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

```
### Creating a chart of all the sentiments across several Austen books

```{r}
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")

```
### Creating a tibble of words in "Pride & Prejudice"

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

### Creating a sentiment analysis of Austen books across NRC, AFINN and Bing
```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

```

```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```
### Creating a count of NRC positive and negative sentiment 
```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)
```
### Creating a count of Bing positive and negative sentiment 
```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```
### Getting Counts of Bing words with their sentimentsnt

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
```
### Chart of the most common word separted by sentiment 

```{r}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```
### Example of creating a custom dictionary
```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```
### Creating a word cloud
```{r}
tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```
### A word cloud divded by Positive and Negative sentiment

```{r}
tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```
### Creating a tibble of sentences 

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
```

```{r}
p_and_p_sentences$sentence[2]

```

### Splitting into tokens using a regex pattern. We could use this, for example, to split the text of Jane Austen’s novels into a data frame by chapter.
```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```
### Creating a Bing list of negative sentiments

```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")
```
### Summarizing the tibble by each chapter
```{r}
wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())
```
### Finding the number of negative words in each chapter and divide by the total words in each chapter
```{r}
tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```
### Creating a Sentiment analysis of State of the Republic Speach by Nelson Mandela in 1996 using the SentimentAnalysis R package 


```{r}
url<-"https://raw.githubusercontent.com/greerda/Data607/main/1996_Mandela.txt"

file_content <- readLines(url)
tokens <- tibble(text = file_content)
cleaned_tibble <- tokens %>% filter(!is.na(text) & text != "")
wordvector<- cleaned_tibble$text
sentiment <- analyzeSentiment(wordvector)
direction_sentiment <- convertToDirection(sentiment$SentimentQDAP)
sentiment_mapping <- c("negative" = -1, "neutral" = 0, "positive" = 1)
response <- sapply(direction_sentiment, function(category) sentiment_mapping[category])
compareToResponse(sentiment, response)


```
### Creating a chart of positive and negative word count 

```{r}
sentiment_categories <- data.frame(category = direction_sentiment)

ggplot(data = sentiment_categories) +
  geom_bar(aes(x = category, fill = category)) +
  ggtitle("Sentiment Category Counts") +
  xlab("Category") +
  ylab("Count") +
  theme(
    plot.title = element_text(hjust=0.5),
    legend.position = "none"
  ) +
  coord_flip()
```

### Creating a word cloud of  of positive and negative words from Nelson Mandela speach.
```{r}
wwordvector<- cleaned_tibble$text
words_list <- unlist(strsplit(wordvector, " "))

sentiment <- analyzeSentiment(words_list)
sentiment$SentimentQDAP
convertToBinaryResponse(sentiment)$SentimentGI
direction_sentiment <- convertToDirection(sentiment$SentimentQDAP)
sentiment_mapping <- c("negative" = -1, "neutral" = 0, "positive" = 1)
response <- sapply(direction_sentiment, function(category) sentiment_mapping[category])
compareToResponse(sentiment, response) 

word_labels <- sapply(direction_sentiment, function(category) {
  sentiment_mapping[category]
})

word_data <- data.frame(Word = words_list, Sentiment = word_labels) %>%
  mutate(Sentiment = ifelse(Sentiment == 0, "negative", "positive")) %>%na.omit(Sentiment)

word_data %>%
  count(Word, Sentiment, sort = TRUE) %>%
  acast(Word ~ Sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```
```


